---
title: "Alex Singleton"
params:
   address: University of Liverpool, Department of Geography and Planning, Roxby Building, Liverpool, L69 7ZT
   email: alex.singleton@liverpool.ac.uk
   telephone: "+44 (0)151 7943108"
   website: www.alex-singleton.com
   
---

<!--html_preserve-->
<style type="text/css">
ul.CV {
list-style-type: none;
}
</style>
<!--/html_preserve-->

```{r echo=FALSE,warning=FALSE,message=FALSE}
library(RefManageR)
```

<ul class="CV">
<li><i class="fa fa-home" aria-hidden="true"></i>  `r paste(" ",params$address)`</li>
<li><i class="fa fa-envelope-o" aria-hidden="true"></i>  `r paste(" ",params$email)`</li>
<li><i class="fa fa-mobile" aria-hidden="true"></i>  `r paste(" ",params$telephone)`</li>
<li><i class="fa fa-globe" aria-hidden="true"></i>  [`r paste(" ",params$website)`](`r params$website`)</li>
<li><i class="fa fa-clock-o" aria-hidden="true"></i> Last Updated: `r paste(" ",format(Sys.Date(), "%B %d, %Y"))`</li>
</ul>


*************

# Education

* 2003 - 07
    + PhD in Geography (Examined Dec. 2007) - University College London, UK.
* 2000 - 03
    + BSc (1st class Hons) in Geography - University of Manchester, UK.

*************

# Awards Honors and Distinctions

* 2016
    + Visiting Professor Ritsumeikan University, Japan.
    + Europe, Middle East and Africa Editor - Transactions in GIS.
* 2014
    + Chair of the Royal Geographical Society (with IBG) GIScience Research Group.
    + Associate Editor of the Spatial Analytics Book Series (Sage).
* 2013
    + Honorary Visiting Researcher Ritsumeikan University, Japan.
* 2012
    + Founding Editor in Chief of Regional Studies, Regional Science.
    + Book Review Editor for Environment and Planning B.
    + Chair GIS Research UK (GISRUK) 2013 Annual Conference.
* 2011
    + Shortlisted for Scopus Young Researcher Award (Social Science).
    + Royal Geographical Society Gildea Award.
* 2010
    + Chair of the Output Area Classification User Group at the Royal Statistical Society.
* 2009
    + Winner GIS Research UK Annual Conference Mashup Competition.
* 2008
    + Chair of the Royal Geographical Society Quantitative Methods Research Group.
* 2006
    + Winner GIS Research UK Annual Conference Poster Competition.
* 2004
    + Winner UCL Graduate School Poster Competition.
* 2003
    + Winner of Manchester Geographical Association Dissertation Prize.

*************

# Research Interests

My research has extended a tradition of area classification within Geography where I have developed an empirically informed critique of the ways in which geodemographic methods can be refined for effective yet ethical use in public resource allocation applications. This research has developed from substantive interests around the social, spatial and temporal dimensions of access inequalities in Higher Education, digital exclusion, aspects of retail and school commuting behaviours. My most recent work has focused on understanding how both fast and slow human dynamics  can be understood within urban systems and their hierarchies, developing new methods of mapping these geographies at fine spatio-temproal scale. More generally I am interested in how geographic data science can be utilised in activities related to the public understanding of science, and I am also an advocate of open source data, software and publishing.

*************

# Academic Roles

* 2015 - 
    + Professor - University of Liverpool, Department of Geography and Planning.
* 2013 - 14
    + Reader - University of Liverpool, Department of Geography and Planning.
* 2010 - 13
    + Lecturer - University of Liverpool, Department of Geography and Planning.
* 2005 - 2010
    + Research Fellow - University College London, Department of Geography.
* 2003-2005
    + Research Assistant - University College London, Department of Geography.

*************

# Research Income and Grants

```{r echo=FALSE}
grants <- read.csv("grants.csv")
```

This following section provides a summary of research income to date which totals `r paste0("£",format(sum(grants$Amount), big.mark=","))`, with `r paste0("£",format(sum(grants[grants$PI_CO_I == "PI",]$Amount), big.mark=","))` as Principal Investigator (PI) and `r paste0("£",format(sum(grants[grants$PI_CO_I == "Co-I",]$Amount), big.mark=","))` as Co-Investigator (Co-I).


```{r echo=FALSE, results="asis",comment=FALSE}

years <- unique(grants$Year)[order(unique(grants$Year),decreasing = TRUE)] #Get year list

for (year in years){#years

  cat("*",year,"\n")
  
  tmp_grant <- grants[grants$Year == year,]
  
for (i in 1:nrow(tmp_grant)){#line printing loop
  cat(paste0("\t"))
  cat("+",paste0("_",tmp_grant[i,]$Funder,"_")," ",paste0(tmp_grant[i,]$Title),paste0("(",tmp_grant[i,]$PI_CO_I,")"),paste0("**£",format(sum(tmp_grant[i,]$Amount), big.mark=",")),"**","\n")
  
}#line print

}#years

```

#Books

```{r echo=FALSE, results="asis",comment=FALSE}
file_books <- ReadBib("books.bib")
dates <- unlist(unique(file_books$year))[order(unlist(unique(file_books$year)),decreasing = TRUE)]
#Prints recerences
for (date in dates) {
  print(file_books[list(year=date)],.opts = list(style="markdown",bib.style ="authoryear" ,max.names =10,dashed=FALSE))
  cat("\n")
}

```

# Refereed Academic Journals

```{r echo=FALSE, results="asis",comment=FALSE}
file_journals <- ReadBib("journals.bib")
dates <- unlist(unique(file_journals$year))[order(unlist(unique(file_journals$year)),decreasing = TRUE)]

#Prints recerences
for (date in dates) {
  cat(paste0("##",date),"\n")
  print(file_journals[list(year=date)],.opts = list(style="markdown",bib.style ="authoryear" ,max.names =10,dashed=FALSE))
  cat("\n")
}

```

# Book Chapters

```{r echo=FALSE, results="asis",comment=FALSE}
file_book_chpt <- ReadBib("book_chapters.bib")
dates <- unlist(unique(file_book_chpt$year))[order(unlist(unique(file_book_chpt$year)),decreasing = TRUE)]

#Prints recerences
for (date in dates) {
  print(file_book_chpt[list(year=date)],.opts = list(style="markdown",bib.style ="authoryear" ,max.names =10,dashed=FALSE))
  cat("\n")
}

```

# Professional Magazines

```{r echo=FALSE, results="asis",comment=FALSE}
file_mag <- ReadBib("magazines.bib")
dates <- unlist(unique(file_mag$year))[order(unlist(unique(file_mag$year)),decreasing = TRUE)]

#Prints recerences
for (date in dates) {
  print(file_mag[list(year=date)],.opts = list(style="markdown",bib.style ="authoryear" ,max.names =10,dashed=FALSE))
  cat("\n")
}

```


# Participation in Conferences and Professional Meetings
## Conferences

```{r echo=FALSE, results="asis",comment=FALSE}
file_conf <- ReadBib("conf.bib")
dates <- unlist(unique(file_conf$year))[order(unlist(unique(file_conf$year)),decreasing = TRUE)]

#Prints recerences
for (date in dates) {
  print(file_conf[list(year=date)],.opts = list(style="markdown",bib.style ="authoryear" ,max.names =10,dashed=FALSE))
  cat("\n")
}

```


## Invited Presentations at Research and Professional Seminars, Conferences and Workshops


```{r echo=FALSE, results="asis",comment=FALSE}
library(RefManageR)
file_invited <- as.data.frame(ReadBib("invited.bib"))
file_invited$title <- gsub("\\{|\\}","",file_invited$title)# Remove {}
file_invited$address <- gsub("\\{|\\}","",file_invited$address)# Remove {}

dates <- unlist(unique(file_invited$year))[order(unlist(unique(file_invited$year)),decreasing = TRUE)]

#Prints references - slightly different format as converted to data frame for formating
for (date in dates) {
  cat(paste0("##",date),"\n")
  
  tmp <- file_invited[file_invited$year==date,]
  
  for (i in 1:nrow(tmp)) {
  cat(paste0("_",tmp[i,]$title,"._ "),paste0(tmp[i,]$address,".","\n"))
  cat("\n")  
    
  }
}

```

# Professional and Administrative Activities
* Europe, Middle East and Africa Editor - Transactions in GIS (2015- ).
* Editor in Chief - Regional Studies and Regional Science (First Open Access journal from Taylor and Francis) (2013- )
* Book Reviews Editor - Environment and Planning B (2012 - 15).
    
# Administrative Experience
* Course director - MSc Applied Geographical Information Science (2014-16). 
* Management of Department of Geography and Planning website (2012-13).
* REF internal reviewer (2012-13).
* Undergraduate admissions (strategic)  (2012-13).
* Year tutor (2011-12).
* REF internal reviewer (2011-12).
* Staff student liaison committee (2010-11).

# External Examination Experience
* University of Warwick: MSc in Big Data and Digital Futures, and MSc in Urban Informatics and Analytics (2015- ).
* University College London: MSc Geospatial Analysis (2016- ).

# Organisational Roles within Professional Societies
* Chair of the GIScience Research Group at the Royal Geographical Society with the Institute of British Geographers (2014–15).
* Chair of the Output Area Classification User Group at the Royal Statistical Society (2010-14).
* Chair of the Quantitative Methods Research Group at the Royal Geographical Society with the Institute of British Geographers (2008-11).


# Membership of Professional Societies
* Fellow of the Royal Geographical Society (2003 -).
* Member of the Association of American Geographers (2003 -).
* Member of the Regional Studies Association (2010 - ).

# Academic and Professional Duties
Referee on articles submitted to the following international journals: 

Area; Environment and Planning B (Planning and Design); Computers, Environment and Urban Systems; International Journal of Health Geographics: International Journal of Pattern Recognition; Papers in Regional Science; Applied Spatial Analysis and Policy; Environmental Modelling; International Journal of Geographic Information Science; Educational Research Journal; URISA Journal; Journal of Maps; Journal of Urban Technology; PLOS One. 

# Conference and Seminar Organisation

* 2016
    + OSGIS 2016 Programme Committee.
* 2014
    + OSGIS 2014 Programme Committee.
* 2013
    + GIS Research UK Conference Chair (http://liverpool.gisruk.org).
    + GIS Ostrava 2013 Member of the Scientific Committee, Czech Republic
* 2012
    + OSGIS 2012 Programme Committee.
    + Session Organiser: New approaches in quantitative social geography: new data, new methods, new insights. Association of American Geographers Conference, New York.
* 2010
    + Making Efficiency Saving with OAC by Understanding Total Place. Output Area Classification User Conference, London.
    + Session Convenor The Spatial Dimensions of Health. Royal Geographical Society Annual International Conference, London.
* 2009
    + Session Convenor New Urban Geography: Evolving Area Classification for Socio-Spatial Generalisation. Royal Geographical Society Annual International Conference, Manchester, 26-28 August.
    + Organising committee for the 2010 GIS Research UK Conference. London, University College London.
    + Conference Organiser - Output Area Classification User Conference.
* 2006
    + Co-Organiser, Spatial Literacy in Teaching: Communicating GIScience, London.
    + Co-Organiser, Spatial Literacy, Student Recruitment and Widening Participation, London.
    
# Impact and Knowledge Transfer
For each research project I have been involved with, a range of academic publications have been produced. In addition to these scholarly outputs, each project is typically accompanied with a website aiming to disseminate the findings more broadly and where possible to engage a wide public audience. These websites take a variety of different formats and are briefly detailed in the following section with evidence for their success measured in terms of press coverage.

## Software
* gd (http://rpubs.com/alexsingleton/gd) - An R package for building, testing and visualisation of geodemographic classification.
* OAC Coder (http://www.publicprofiler.org/resources.php) - Postcode profiling tool for the ONS Output Area Classification (downloaded over 700 times).
## Websites
I have developed and/or manage a range successful websites created as part of activities associated with research grants. These websites have wide ranging purposes, including pedagogic resource dissemination, public participation in GIS or impact generation. Many of these sites have captured a wide and international audience. For example, the worldnames (worldnames.publicprofler.org) website has received over 2 million unique users from 200 countries since August 30 2008.

* www.policestreets.co.uk - a new visualisation of crime data.
* rpubs.com/alexsingleton/ - R tutorials.
* www.educationprofiler.org - School and catchment profiles with Google Maps.
* gbnames.publicprofiler.org - 1881 and 1998 GB surname mapping website.
* worldnames.publicprofler.org - world surname mapping website. 
* esociety.publicprofler.org - e-Society profiler website (search by postcode).
* www.onomap.org - name classification website.
* www.publicprofiler.org - neighbourhood profiler website (search by postcode).
* www.areaclassification.org.uk - geodemographic news and teaching materials.
* www.alex-singleton.com - personal research blog.

# Media Exposure
My own research and/or the projects that I have worked on or managed have gained attention through the following media features:

* 2013
    + Mapping the Census: How One Man Produced a Library for All. Guardian: (http://goo.gl/HnazXF).
* 2011
    + The Geography of Poverty. Socialist Review: (http://goo.gl/NoQjx7).
    + 10/02 - UK Crime Map Commentary. BBC Radio Berkshire.
    + 01/05 - The End of the Census? Geographical Magazine.
    + 27/9 - Why do people riot? Current Biology: (http://goo.gl/Uto0nA).
    + 16/8 - England Riots: Was Poverty a Factor. Guardian: (http://goo.gl/lPx7u8).
    + 18/8 - England Rioters: Young Poor and Unemployed: (http://goo.gl/2NHloc).
    + 11/11 - TUC Briefing: Responding to the Riots.
* 2008
    + 10/04 - Nestoria Interview - Alex Singleton - UCLs Centre for Advanced Spatial Analysis. Nestoria Weblog: (http://goo.gl/BZ1ugq).
    + 01/11 - A Multicultural Atlas of London. ESRC Britain in 2009 Magazine, p 100.
    + 01/11 - Using Geodemographics in Public Service Delivery. ESRC Britain in 2009 Magazine, pp 28.
    + 01/11 - Tracing Family Histories goes Global. ESRC Britain in 2009 Magazine, pp 111.
    + 30/08 - Website Maps Surnames Worldwide. BBC News Online: (http://goo.gl/dIKw0q).
    + 30/08 - Putting you on the Map: the Website that Pinpoints where your Name is in the World. The Independent.
    + 30/08 - Global Surname Website Launched. Channel 4 News.
    + 30/08 - What’s in a Name? A Great Deal, Say Researchers. The Telegraph: (http://goo.gl/r37HMn).
    + 30/08 - Website Pinpoints Where in the World Your Name is. New Zealand Herald: (http://goo.gl/MpxZOm)
    
* 2006
    + 10/04 - 08/08 - Britain’s Digital Tribes Revealed. BBC News Online: (http://goo.gl/SDlnAv).
    + 15/01 - Mr Smith? From Lerwick, I presume, The Observer: (http://goo.gl/urZbPV) 16/01 - Whats in a Surname? Channel 4.
    + 21/01 - My, You’ve Come a Long Way Without Even Leaving Home. The Times
    + 21/01 - The Internet Route to the Roots of Your Surname. The Daily Mail. 25/01 - Miles Mendoza Website of the Day. Radio 2 (http://goo.gl/r3YRgA).
    + 04/06 - Name check reveals a new pecking order. The Sunday Times.
    + 05/06 - How a Cadbury Outranks a Blair or a Beckham. The Daily Mail.
    + 09/06 - Have I Got News For You. BBC Two.
    + 31/08 - End of the Innuendo. The Guardian (http://goo.gl/Mt5rnk).
    + 31/08 - How British names conquered the world. The Telegraph (http://goo.gl/BKAs97).
    
# Other Education and Qualifications
* NVQ Level 4 in Management (part time) (2003-05).
* NVQ Level 3 in Supervisory Management (part time) (1998-99). 
* A Levels in Biology (B), Business (B), Geography (A) (1996-98)

# Other Work Experience
* Researcher - Open University: Knowledge Media Institute (Summer Work) (2003). 
* Web Designer - BDH TBWA New Media Department (Summer Work) (2002).
* Marketing and Information Technology Assistant - Cascade Consulting (Summer Work) (2002).
* Researcher - Open University: Centre for * New Media (Summer Work) (2001).
* Researcher - Open University: Knowledge Media Institute (Summer Work) (2000).
* Web Developer and Webmaster - Denbigh School (Gap Year) (1999).
* Recruitment Assistant - Alenia Marconi (Gap Year- Year in Industry Scheme) (1998).
    